### 2020-06-04 writeup/intro.tex
intro wip

Understanding positive vs. negative transfer requires developing generalization bounds that scale tightly with properties of the data such as sample size.
This presents a technical challenge in the multi-task setting due to the presence of different task features, even for two tasks.
For Rademacher complexity or VC-based techniques, the generalization error scales down as the sample sizes of all tasks increase, when applied to the multi-task setting \cite{B00,AZ05,M06,MPR16,WZR20}.
Without a tight lower bound for the multi-task setting, it is unclear how to compare its performance to single-task learning.
Furthermore, developing a better understanding of how multi-task learning performs in terms of properties of the data such as sample size is of practical importance, since a major application of multi-task learning is for settings where a particular task has limited amount of labeled data \cite{RH19}.
%For example, the sample sizes of all tasks are often assumed to be equal \cite{B00,LPTV09,LPVT11}.
%On the other hand, uneven sample sizes (or dominating tasks) have been empirically observed to cause negative transfer \cite{YKGLHF20}.
%The benefit of learning multi-task representations has also been studied for certain half-spaces \cite{} and sparse regression \cite{}.
%When all tasks are sufficiently similar, adding more labeled data improves the generalization performance for predicting a particular task \cite{WZR20}.

%\textbf{Setup and Main Results.}
In this work, we apply recent development in the random matrix theory literature to show how to determine positive vs. negative transfer in the setting of high-dimensional linear regression \cite{HMRT19,BLLT20}.
This setting allows us to rigorously quantify three components, defined as follows.
For simplicity, we consider two tasks with first being the source and the second being the target, but our formulation applies more generally.
\squishlist
	\item \textbf{Task model distance:} Assume that each task follows a linear model with parameters $\beta_i\in\real^p$, for $i=1, 2$.
	Hence, $\norm{\beta_1 - \beta_2}$ measure the distance of the two tasks.
	\item \textbf{Sample size:} Let $n_i = \rho_i \cdot p$ denote the sample size of task $i$, for a fixed value $\rho_i > 1$ that does not grow with $p$, for any $1\le i\le t$.
	Hence, the source/target sample ratio is measured by $n_1 / n_2 = \rho_1 / \rho_2$.
	Importantly, we assume that $\rho_2$ is a small constant (say 2), to capture the need for adding more labeled data.
	\item \textbf{Covariate shift:} We assume that for every task, its features are random vectors with covariance matrix $\Sigma_i\in\real^{p\times p}$.
	%$x = \Sigma_i^{1/2}z$, where $z\in\real^p$ consists of i.i.d. entries with mean zero and unit variance, and is a positive semidefinite matrix.
	We measure covariate shift by $\Sigma_i^{1/2}\Sigma_j^{-1/2}$.
\squishend
To motivate our study, Figure \ref{fig_model_shift_phasetrans} shows three examples for two linear regression tasks solved using a shared linear layer for both tasks and an output layer for each task.
We observe that negative transfer occurs as (a) \textit{task similarity}: tasks become more different; (b) \textit{data size}: source/target data size increases.
%\textbf{Task similarity:}
%\textbf{Data sizes:}
%\textbf{Covariate shift:}
Furthermore, MTL performance is negatively affected when (c) \textit{covariate shift}: the covariance matrices of the two tasks become more different.
The observations highlight the need to develop generalization bounds that scale tightly with properties of multiple tasks data.
% that only contains limited amount of labeled data.
%Following Hastie et al.  and Bartlett et al. \cite{},

>>>>

Understanding positive vs. negative transfer requires developing generalization bounds that scale tightly with properties of the data such as sample size.
This presents a technical challenge in the multi-task setting due to the presence of different task features, even for two tasks.
For Rademacher complexity or VC-based techniques, the generalization error scales down as the sample sizes of all tasks increase, when applied to the multi-task setting \cite{B00,AZ05,M06,MPR16,WZR20}.
Without a tight lower bound for the multi-task setting, it is unclear how to compare its performance to single-task learning.
Furthermore, developing a better understanding of how multi-task learning performs in terms of properties of the data is of practical importance, since a major application of multi-task learning is for settings where a particular task has limited amount of labeled data \cite{RH19}.
%For example, the sample sizes of all tasks are often assumed to be equal \cite{B00,LPTV09,LPVT11}.
%On the other hand, uneven sample sizes (or dominating tasks) have been empirically observed to cause negative transfer \cite{YKGLHF20}.
%The benefit of learning multi-task representations has also been studied for certain half-spaces \cite{} and sparse regression \cite{}.
%When all tasks are sufficiently similar, adding more labeled data improves the generalization performance for predicting a particular task \cite{WZR20}.

%\textbf{Setup and Main Results.}
In this work, we apply recent development in the random matrix theory literature \cite{erdos2017dynamical,isotropic,Anisotropic} to show how to determine positive vs. negative transfer in the setting of high-dimensional linear regression \cite{HMRT19,BLLT20}.
This setting allows us to measure the difference of task data by \textit{task similarity}, \textit{sample size} and \textit{covariate shift} precisely.
We define each component for two tasks and our formulation applies more generally.
We call the first task a source task and the second a target task.
\squishlist
	\item \textbf{Task model distance:} Assume that each task follows a linear model with parameters $\beta_i\in\real^p$, for $i=1, 2$.
	Hence, $\norm{\beta_1 - \beta_2}$ measure the distance of the two tasks.
	\item \textbf{Sample size:} Let $n_i = \rho_i \cdot p$ denote the sample size of task $i$, for a fixed value $\rho_i > 1$ that does not grow with $p$, for any $1\le i\le t$.
	Hence, the source/target sample ratio is measured by $n_1 / n_2 = \rho_1 / \rho_2$.
	Importantly, we assume that $\rho_2$ is a small constant (say 2), to capture the need for adding more labeled data.
	\item \textbf{Covariate shift:} We assume that for every task, its features are random vectors with covariance matrix $\Sigma_i\in\real^{p\times p}$.
	%$x = \Sigma_i^{1/2}z$, where $z\in\real^p$ consists of i.i.d. entries with mean zero and unit variance, and is a positive semidefinite matrix.
	We measure covariate shift by $\Sigma_i^{1/2}\Sigma_j^{-1/2}$.
\squishend
To motivate our study, Figure \ref{fig_model_shift_phasetrans} shows three examples for two linear regression tasks solved using a shared linear layer for both tasks and an output layer for each task.
We observe that negative transfer occurs as (a) \textit{task similarity}: tasks become more different; (b) \textit{data size}: source/target data size increases.
%\textbf{Task similarity:}
%\textbf{Data sizes:}
%\textbf{Covariate shift:}
Furthermore, MTL performance is negatively affected when (c) \textit{covariate shift}: the covariance matrices of the two tasks become more different.
The observations highlight the need to develop generalization bounds that scale tightly with properties of multiple tasks data.
% that only contains limited amount of labeled data.
%Following Hastie et al.  and Bartlett et al. \cite{},
